---
title: "SocKult analysis"
author: "Annamaya Halskov-Jensen"
date: "14/5/2019"
output: pdf_document
---

LOADING AND PREPARATION
```{r}

setwd("/Users/annamayahalskov-jensen/Documents/Cognitive Science/4th semester/Social and cultural dynamics in cognition/SocKult Exam")
library(pacman)
p_load("coda", "mvtnorm", "devtools", "rstan", "rethinking", "tidyverse", "dfcrm", "brms", "RColorBrewer", "MCMCvis")

data <- read.csv("full_set.csv")
data_test <- data %>% select(ID, condition, steps)

data_test$condition <- as.character(data_test$condition)
data_test$condition[data_test$condition == "cooperative"] <- "0"
data_test$condition[data_test$condition == "competitive"] <- "1"
data_test$condition <- as.factor(data_test$condition)

#Standardize predictors to a mean of 0 and sd of 1
data_test$steps.s = (data_test$steps - mean(data_test$steps, na.rm = T))/ sd(data_test$steps)

```


PLOTS
```{r}
ggplot(data, aes(steps)) + 
  geom_density(aes(fill = condition, colour = condition), alpha = 0.3) + 
  facet_wrap(~condition)

ggplot(data, aes(steps)) + 
  geom_density(aes(fill = competence, colour = competence), alpha = 0.3) + 
  facet_wrap(~competence)

ggplot(data, aes(tom, steps)) +
  geom_point(aes(colour = condition))

ggplot(data, aes(competence, steps)) +
  geom_point(aes(colour = condition))
```


PREDICTIVE PRIOR CHECK(S)
```{r}
m0_prior <- brm(steps ~ 1 + condition, family = gaussian, data, prior = prior_20, sample_prior = "only")
m4.1_prior <- brm(steps ~ 1 + condition + tom + competence, family = gaussian, data, prior = prior_20, sample_prior = "only")
m1.3_prior <- brm(steps ~ 1 + tom + competence, family = gaussian, data, prior = prior, sample_prior = "only")
m3.1_prior <- brm(steps ~ 1 + condition * tom + competence, family = gaussian, data, prior = prior, sample_prior = "only")
 
pp_check(m0_prior, nsample = 100)
pp_check(m4.1_prior, nsample = 100)
 
```


0.0: INTERCEPT ONLY
```{r}  
# Getting the priors
get_prior(steps ~ 1, data, family = gaussian)
 
# Defining priors used
prior0 = c(
  prior(normal(20, 5), class = 'Intercept'), 
  prior(normal(0, 10), class = 'sigma'))

# modelling 
m0.0 <- brm(steps ~ 1, family = gaussian, data, prior = prior0)
m0.0

```


0: CONDITION ONLY
```{r} 
# Getting the priors
get_prior(steps ~ 1 + condition, data, family = gaussian)

# Defining priors used
prior_20 = c(
  prior(normal(20, 10), class = 'Intercept'), 
  prior(normal(0, 10), class = 'b'), 
  prior(normal(0, 10), class = 'sigma'))

prior_10 = c(
  prior(normal(10, 10), class = 'Intercept'), 
  prior(normal(0, 10), class = 'b'), 
  prior(normal(0, 10), class = 'sigma'))

prior_new = c(
  prior(normal(20, 5), class = 'Intercept'), 
  prior(normal(0, 10), class = 'b'), 
  prior(normal(0, 10), class = 'sigma'))

# modelling 
m0_prior <- brm(steps ~ 1 + condition, family = gaussian, data, prior = prior_new, sample_prior = "only")
m0 <- brm(steps ~ 1 + condition, family = gaussian, data, prior = prior_new)
m0

# If set to "only", samples are drawn solely from the priors ignoring the likelihood, which allows among others to generate samples from the prior predictive distribution. In this case, all parameters must have proper priors.

# Checking priors
pp_check(m0_prior, nsample = 100)

# MAP
flist <-  alist(
    steps ~ dnorm(mu , sigma),
    mu <- a + b * condition,
        a ~ dnorm(10, 10),
        b ~ dnorm(0, 10),
        sigma ~ dcauchy(0, 2))

m.m0 <- rethinking::map2stan(flist, data = data)
precis(m.m0)
precis_plot(precis(m.m0))
```


WAIC COMPARISON MODEL 0.0 + 0
```{r}
# compute and save the WAIC information for the next three models 
m0 <- add_criterion(m0, "waic") 
m0.0 <- add_criterion(m0.0, "waic")

# compare the WAIC estimates
w <- loo_compare(m0, m0.0, criterion = "waic")
print(w, simplify = F)

#Calculating weights
model_weights(m0.b, m1.1.b, m1.2.b, weights = "waic") %>% 
  round(digits = 2)
```


1.1: CONDITION + TOM
```{r}
# modelling in brms
m1.1_prior <- brm(m1, family = gaussian, data, prior = prior, sample_prior = "only")
m1.1 <- brm(steps ~ 1 + condition + tom, family = gaussian, data, prior = prior_20)
m1.1

# Checking priors
pp_check(m1_prior, nsample = 100)
```


1.2: CONDITION + COMPETENCE
```{r}
# modelling in brms
m1.2_prior <- brm(m2, family = gaussian, data, prior = prior, sample_prior = "only")
m1.2 <- brm(steps ~ 1 + condition + competence, family = gaussian, data, prior = prior_20)
m1.2
 
plot(m2.b)

# Checking priors
pp_check(m2_prior, nsample = 100)
```


1.3: TOM + COMPETENCE
```{r}
# modelling in brms 
m1.3_prior <- brm(steps ~ 1 + tom + competence, family = gaussian, data, prior = prior, sample_prior = "only")
m1.3 <- brm(steps ~ 1 + tom + competence, family = gaussian, data, prior = prior_new)
m1.3

plot(m1.3.b) 

# Checking priors
pp_check(m4.2_prior, nsample = 100)

```


WAIC COMPARISON MODEL 0 + 1.1 + 1.2 + 1.3
```{r} 
# compute and save the WAIC information for the next three models 
m0 <- add_criterion(m0, "waic")
m1.1 <- add_criterion(m1.1, "waic")
m1.2 <- add_criterion(m1.2, "waic")
m1.3 <- add_criterion(m1.3, "waic")

# compare the WAIC estimates
w <- loo_compare(m0, m1.1, m1.2, m1.3, criterion = "waic")
print(w, simplify = F)

#Calculating weights
model_weights(m0.b, m1.1.b, m1.2.b, weights = "waic") %>% 
  round(digits = 2)
```


2.1: CONDITION * TOM
```{r}
# modelling in brms
m2.1_prior <- brm(steps ~ 1 + condition * tom, family = gaussian, data, prior = prior, sample_prior = "only")
m2.1 <- brm(steps ~ 1 + condition * tom, family = gaussian, data, prior = prior_20)
m2.1

plot(m2.1.b)

# Checking priors
pp_check(m2.1_prior, nsample = 100)

# MAP
flist <-  alist(
    steps ~ dnorm(mu , sigma),
    mu <- a + bc * condition + bcc * competence,
        a ~ dnorm(50, 10),
        bc ~ dnorm(20, 10),
        bcc ~ dnorm(10, 10),
        sigma ~ dunif(0, 10))

m.m2 <- rethinking::map(flist, data = data)
```


WAIC COMPARISON MODEL 1.1 + 2.1
```{r}
# compute and save the WAIC information for the next three models
m2.1 <- add_criterion(m2.1, "waic")

# compare the WAIC estimates
w <- loo_compare(m1.1, m2.1, criterion = "waic")
print(w, simplify = F)

#Calculating weights
model_weights(m2.1.b, m2.1.b, weights = "waic") %>% 
  round(digits = 2)  
```


3.1: CONDITION * TOM + COMPETENCE
```{r}
# modelling in brms
m3.1_prior <- brm(steps ~ 1 + condition * tom + competence, family = gaussian, data, prior = prior, sample_prior = "only")
m3.1 <- brm(steps ~ 1 + condition * tom + competence, family = gaussian, data, prior = prior_new)
m3.1

plot(m3.1.b)

# checking priors
pp_check(m3.1_prior, nsample = 100)


# marginal effects
# marginal, m3.1
that <- brms::marginal_effects(m3.1)
plot(that, plot = F)[[4]] +
  scale_color_brewer(type = "qual", palette = "Set1") +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  labs(title = "Marginal effects of ToM & condition", x = "ToM", y = "Steps from $100") +
  theme_minimal()

```


3.2: CONDITION * TOM * COMPETENCE
```{r}
# modelling in brms
m3.2_prior <- brm(steps ~ 1 + condition * tom * competence, family = gaussian, data, prior = prior, sample_prior = "only")
m3.2 <- brm(steps ~ 1 + condition * tom * competence, family = gaussian, data, prior = prior_20)
m3.2

plot(m3.2.b)
 
# Checking priors
pp_check(m3.1_prior, nsample = 100) 
```


WAIC COMPARISON MODEL 2.1 + 3.1 + 3.2
```{r} 
# compute and save the WAIC information for the next three models 
m3.1 <- add_criterion(m3.1, "waic")
m3.2 <- add_criterion(m3.2, "waic")

# compare the WAIC estimates
w <- loo_compare(m3.1, m3.2, criterion = "waic")
print(w, simplify = F)


w <- loo_compare(m2.1.b, m0.b, criterion = "waic")
print(w, simplify = F)

#Calculating weights
model_weights(m2.1.b, m3.1.b, m3.1.b, weights = "waic") %>% 
  round(digits = 2) 
```


4.0: TOM * COMPETENCE
```{r}
# modelling in brms
m4.0_prior <- brm(steps ~ 1 + tom * competence, family = gaussian, data, prior = prior, sample_prior = "only")
m4.0 <- brm(steps ~ 1 + tom * competence, family = gaussian, data, prior = prior_20)
m4.0

plot(m4.0.b)

# Checking priors
pp_check(m4.0_prior, nsample = 100)

```


4.1: CONDITION + TOM + COMPETENCE
```{r} 
# modelling in brms
m4.1_prior <- brm(steps ~ 1 + condition + tom + competence, family = gaussian, data, prior = prior, sample_prior = "only")
m4.1 <- brm(steps ~ 1 + condition + tom + competence, family = gaussian, data, prior = prior_new)
m4.1 
 
# investigating priors
m4.1_10 <- brm(steps ~ 1 + condition + tom + competence, family = gaussian, data, prior = prior_10)
m4.1_20 <- brm(steps ~ 1 + condition + tom + competence, family = gaussian, data, prior = prior_20)


plot(m4.0.b) 

# Checking priors
p0 <- pp_check(m0, nsample = 20)
p4.1 <- pp_check(m4.1, nsample = 20)

plot_grid(p0, p4.1)


```


4.2 CONDITION + TOM * COMPETENCE
```{r}
# modelling in brms
m4.2_prior <- brm(steps ~ 1 + condition + tom * competence, family = gaussian, data, prior = prior, sample_prior = "only")
m4.2 <- brm(steps ~ 1 + condition + tom * competence, family = gaussian, data, prior = prior_20)
 m4.2
 
plot(m4.2.b) 

# Checking priors
pp_check(m4.2_prior, nsample = 100)
```


WAIC COMPARISON MODEL 4.0 + 4.1 + 4.2
```{r}  
# compute and save the WAIC information for the next three models
m4.0 <- add_criterion(m4.0, "waic")
m4.1 <- add_criterion(m4.1, "waic")
m4.2 <- add_criterion(m4.2, "waic")

# compare the WAIC estimates
w <- loo_compare(m4.0.b, m4.1.b, m4.2.b, criterion = "waic")
print(w, simplify = F)


w <- loo_compare(m2.1.b, m0.b, criterion = "waic")
print(w, simplify = F)

#Calculating weights
model_weights(m2.1.b, m3.1.b, m3.1.b, weights = "waic") %>% 
  round(digits = 2) 
```


4.3: TOM
```{r}
# modelling in brms
m4.3_prior <- brm(steps ~ 1 + tom, family = gaussian, data, prior = prior, sample_prior = "only")
m4.3 <- brm(steps ~ 1 + tom, family = gaussian, data, prior = prior_20)
m4.3

plot(m4.2.b)

# Checking priors
pp_check(m4.2_prior, nsample = 100)

```


4.4: COMPETENCE
```{r}
# modelling in brms
m4.4_prior <- brm(steps ~ 1 + competence, family = gaussian, data, prior = prior, sample_prior = "only")
m4.4 <- brm(steps ~ 1 + competence, family = gaussian, data, prior = prior_20)
m4.4

plot(m4.4.b)

# Checking priors
pp_check(m4.2_prior, nsample = 100)

```


WAIC COMPARISON MODEL 4.0 + 4.1 + 4.2
```{r}
# compute and save the WAIC information for the next three models
m4.2 <- add_criterion(m4.3, "waic")
m4.3 <- add_criterion(m4.3, "waic")
m4.4 <- add_criterion(m4.4, "waic")

# compare the WAIC estimates
w <- loo_compare(m0.b, m4.3.b, m4.4.b, criterion = "waic")
print(w, simplify = F)

all <- loo_compare(m0.0, m0, m1.1, m1.2, m1.3, m2.1, m3.1, m3.2, m4.0, m4.1, m4.2, m4.3, m4.4, criterion = "waic")
print(all, simplify = F)

#Calculating weights
model_weights(m2.1.b, m3.1.b, m3.1.b, weights = "waic") %>% 
  round(digits = 2) 
```


5.1: CONDITION + TOM + COMPETENCE, zero inflated poisson
```{r}
get_prior(steps ~ 1 + condition + tom + competence, data, family = zero_inflated_poisson())

prior_2 = c(
  prior(normal(50, 10), class = 'Intercept'), 
  prior(normal(0, 10), class = 'b'), 
  prior(normal(0, 10), class = 'zi'))

# modelling in brms
m5.1_prior <- brm(steps ~ 1 + condition + tom + competence, family = zero_inflated_poisson(), data, prior = prior_2, sample_prior = "only")
m5.1 <- brm(steps ~ 1 + condition + tom + competence, family = zero_inflated_poisson(), data, prior = prior_2)
m5.1

m5.1 <- add_criterion(m5.1, "waic")
 
plot(m5.1) 

# Checking priors
pp_check(m5.1_prior, nsample = 100)

all <- loo_compare(m0, m1.1, m1.2, m1.3, m2.1, m3.1, m3.2, m4.0, m4.1, m4.2, m4.3, m4.4, m5.1, criterion = "waic")
print(all, simplify = F)

```


5.1: CONDITION + TOM + COMPETENCE, zero inflated poisson
```{r}
get_prior(steps ~ 1 + condition + tom + competence, data, family = poisson())

prior_3 = c(
  prior(normal(50, 10), class = 'Intercept'), 
  prior(normal(0, 10), class = 'b'))

# modelling in brms
m5.2_prior <- brm(steps ~ 1 + condition + tom + competence, family = zero_inflated_poisson(), data, prior = prior_2, sample_prior = "only")
m5.2 <- brm(steps ~ 1 + condition + tom + competence, family = poisson(), data, prior = prior_3)
m5.2

m5.2 <- add_criterion(m5.2, "waic")
 
plot(m5.2) 

# Checking priors
pp_check(m5.2_prior, nsample = 100)

few <- loo_compare(m0, m4.1, m3.1, m1.3, criterion = "waic")
print(few, simplify = F)

some <- loo_compare(m0, m0.0, m3.1, m4.1, m1.3, criterion = "waic")
print(some, simplify = F)

all <- loo_compare(m0, m0.0, m1.1, m1.2, m1.3, m2.1, m3.1, m3.2, m4.0, m4.1, m4.2, m4.3, m4.4, criterion = "waic")
print(all, simplify = F)

#Investigating interaction
interaction <- loo_compare(m0, m3.2, m3.1, m4.1, m4.0, m4.2, criterion = "waic")
print(interaction, simplify = F)

```



```{r}
p_load(cowplot)

p1 <- pp_check(m3.1) +
  facet_wrap(~data$condition) +
  ggtitle("PPC for conservative model")
p1

p2 <- pp_check(m4.1) +
  facet_wrap(~data$condition)+
  ggtitle("PPC for metaanalytic priors")
p2

plot_grid(p1, p2)

pred <- predict(m4.1)
pred <- as.data.frame(pred)
pred$actual <- data$steps
pred$condition <- data$condition

p_steps <- ggplot(pred, aes(actual)) + 
  geom_density(aes(fill = condition, colour = condition), alpha = 0.3) + 
  facet_wrap(~condition)

p_pred <- ggplot(pred, aes(Estimate)) + 
  geom_density(aes(fill = condition, colour = condition), alpha = 0.3) + 
  facet_wrap(~condition)

plots4.1 <- plot_grid(p_steps, p_pred)

#Investigating interaction condition * tom
pred3.1 <- predict(m3.1)
pred3.1 <- as.data.frame(pred3.1)
pred3.1$actual <- data$steps
pred3.1$condition <- data$condition

p3.1_steps <- ggplot(pred3.1, aes(actual)) + 
  geom_density(aes(fill = condition, colour = condition), alpha = 0.3) + 
  facet_wrap(~condition)

p3.1_pred <- ggplot(pred3.1, aes(Estimate)) + 
  geom_density(aes(fill = condition, colour = condition), alpha = 0.3) + 
  facet_wrap(~condition)

plots3.1 <- plot_grid(p3.1_steps, p3.1_pred)

p_load(gridExtra, lattice)
grid.arrange(p_steps, p_pred, p3.1_steps, p3.1_pred)




ggplot(pred, aes(actual, Estimate)) +
  geom_smooth(aes(colour = condition), method = "lm") +
  geom_point(aes(colour = condition)) +
  geom_abline(xintercept = 100, yintercept = 100, linetype = "dashed") +
  xlim(0, 100) +
  ylim(0, 100)

ggplot(pred0, aes(actual, Estimate)) +
  geom_smooth(aes(colour = condition), method = "lm") +
   geom_point() +
  geom_abline(xintercept = 100, yintercept = 100, linetype = "dashed") +
  xlim(0, 100) +
  ylim(0, 100)

```


```{r}
# Plotting predictions of m1.3
pred3 <- predict(m1.3)
pred3 <- as.data.frame(pred3)
pred3$actual <- data$steps
pred3$tom <- data$tom
pred3$competence <- data$competence

p3_steps <- ggplot(pred3, aes(actual)) + 
  geom_density(aes(fill = tom, colour = tom), alpha = 0.3)

p3_pred <- ggplot(pred3, aes(Estimate)) + 
  geom_density(aes(fill = competence, colour = competence), alpha = 0.3)

plots3 <- plot_grid(p3_steps, p3_pred)

ggplot(pred3, aes(Estimate, actual)) + 
  geom_point(aes(fill = competence, colour = tom), alpha = 0.3)

ggplot(pred3, aes(actual, Estimate)) +
  geom_smooth(color = "#449999", method = "lm", se = F) +
  geom_point(color = "Dark red") +
  geom_abline(xintercept = 100, yintercept = 100, linetype = "dashed") +
  xlim(0, 100) +
  ylim(0, 100) +
  theme_minimal() +
  labs(title = "Observation vs. Prediction - Model 3", x = "Observed values (steps)", y = "Model predictions (steps)")

```




```{r} 
xlab <- "Trial" 
ylab <- "Measure of Theory of Mind"
num <- 2
getPalette <- colorRampPalette(brewer.pal(8, "Accent"))

ggplot(tom, aes(trial, tom)) +
  geom_point(aes(colour = ID), size = 2) +
  scale_color_manual(values = getPalette(num)) +
  geom_line(aes(colour = ID), size = 1) +
  labs(x = xlab, y = ylab) +
  scale_y_continuous(breaks = c(-10:20)) +
  scale_x_continuous(breaks = c(1:3))
```




zero-inflated models
- poisson
